Carbonate chemistry at Ny-Ålesund: AWIPEV-CO~2~ Project

### 2020-09-04 awipev-CO2.Rmd
more on stratification

# we look at salinity 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth) %>%
  dplyr::mutate(depth = as.factor(ifelse(depth == 1 | depth == 3, 2, depth))) %>% 
  dplyr::filter(depth == 2 | depth == 9) %>% 
  group_by(depth, montht)
summary(aov(tmp$temp_insitu_ctd ~ tmp$depth * tmp$montht))

p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

@@

# we look at temperature 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth) %>%
  dplyr::mutate(depth = as.factor(ifelse(depth == 1 | depth == 3, 2, depth))) %>% 
  dplyr::filter(depth == 2 | depth == 9) %>% 
  group_by(depth, montht)
summary(aov(tmp$temp_insitu_ctd ~ tmp$depth * tmp$montht))

@@

# we look at density 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, temp_insitu_ctd, depth) %>%
  dplyr::mutate(density = seacarb::rho(S = sal_insitu_ctd, T = temp_insitu_ctd, P = 0),
                depth = as.factor(ifelse(depth == 1 | depth == 3, 2, depth))) %>% 
  dplyr::filter(depth == 2 | depth == 9) %>% 
  group_by(depth, montht)
summary(aov(tmp$density ~ tmp$depth * tmp$montht))

p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = density)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Density vs depth", x = "Depth (m)", y= "Salinity") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
```

>>>>

# heatmap
tmp <- d_all %>%
  dplyr::filter(!is.na(depth)) %>%
  dplyr::select(montht, sal_insitu_ctd, depth) %>%
  group_by(montht, depth)
sal_median <- skim(tmp) %>%
  dplyr::select(depth, montht, numeric.p50) %>%
  pivot_wider(names_from = depth, values_from = c(numeric.p50))

plt <- sal_median %>%
  ggplot(aes(
    x = montht,
    y = as.factor(depth),
    fill = median(sal_insitu_ctd)
  ), color = "white") +
  geom_tile() +
  scale_fill_gradientn(colors = viridis(5), breaks = seq(31, 36, by = 1), na.value = "grey") +
  labs(x = "", y = "", fill = "Salinity") +
  scale_x_discrete(position = "top") +
  scale_y_discrete(name = "", limits = rev(levels(as.factor(tmp$depth)))) +
  # scale_y_discrete(name = "", limits = rev(levels(as.factor(tmp$depth))),
  #                  rev(levels(tmp$depth))[c(1,3,5,7,9)]) +
  # coord_equal() +
  theme(legend.position = "top",
        axis.text.y = element_text(size = 13),
        axis.text.x = element_text(size = 13),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 14),
        #legend.key.size = unit(1.5, "cm"),
        legend.key.height = unit(0.8, "cm"),
        legend.key.width  = unit(2, "cm"))
plt

# we look at salinity 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, depth) %>%
  dplyr::mutate(depth = as.factor(ifelse(depth == 1 | depth == 3, 2, depth))) %>% 
  dplyr::filter(depth == 2 | depth == 9) %>% 
  group_by(depth, montht)
sal_range <- skim(tmp) %>% 
  dplyr::select(depth, montht, numeric.p50) %>% 
  pivot_wider(names_from = depth, values_from = c(numeric.p50)) %>% 
  dplyr::rename(median2m = "2",
                median9m = "9") %>% 
  dplyr::mutate(sal_range = median2m - median9m)
kable(sal_range, digits = 2, caption = "Median salinity")

summary(aov(tmp$sal_insitu_ctd ~ tmp$depth * tmp$montht))

p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

@@

# we look at temperature 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth) %>%
  dplyr::mutate(depth = as.factor(ifelse(depth == 1 | depth == 3, 2, depth))) %>% 
  dplyr::filter(depth == 2 | depth == 9) %>% 
  group_by(depth, montht)
temp_range <- skim(tmp) %>% 
  dplyr::select(depth, montht, numeric.p50) %>% 
  pivot_wider(names_from = depth, values_from = c(numeric.p50)) %>% 
  dplyr::rename(median2m = "2",
                median9m = "9") %>% 
  dplyr::mutate(temp_range = median2m - median9m)
kable(temp_range, digits = 2, caption = "Median temperature")
summary(aov(tmp$temp_insitu_ctd ~ tmp$depth * tmp$montht))

@@

# we look at density 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, temp_insitu_ctd, depth) %>%
  dplyr::mutate(density = seacarb::rho(S = sal_insitu_ctd, T = temp_insitu_ctd, P = 0),
                depth = as.factor(ifelse(depth == 1 | depth == 3, 2, depth))) %>% 
  dplyr::filter(depth == 2 | depth == 9) %>% 
  dplyr::select(montht, depth, density) %>% 
  group_by(depth, montht) 
den_range <- skim(tmp) %>% 
  dplyr::select(depth, montht, numeric.p50) %>% 
  pivot_wider(names_from = depth, values_from = c(numeric.p50)) %>% 
  dplyr::rename(median2m = "2",
                median9m = "9") %>% 
  dplyr::mutate(den_range = median2m - median9m)
kable(den_range, digits = 2, caption = "Median density")
summary(aov(tmp$density ~ tmp$depth * tmp$montht))

p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = density)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Density vs depth", x = "Depth (m)", y= "Salinity") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
```
Using the medians because data are missing.

+ Salinity 
+ Temperature is always colder at 0-4 m than below 8 m (0.15 to 2°C).
